---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# localSCR

<!-- badges: start -->
<!-- badges: end -->

The goal of localSCR is to provide user-friendly functions to implement spatially-explicit mark recapture (SECR) models using the nimble package in R. The package currently has functions to 1) assist with defining the state-space grid and extent for a given 2-dimensional or 3-dimensional trap array (i.e., when traps are clustered in space), 2) simulate data under different encounter distributions and other parameters, 3) create habitat masks from either raster data or spatial polygons, 4) provide template SECR models that are easily customizable, and 4) fit and summarize SECR models using nimble with options for parallel processing. Future functionality will include 1) creating realized density surfaces from MCMC output and 2) implementing localized approaches as in Milleret et al. (2019) and Woodruff et al. (2020). The package also uses block updating of x-and y- activity center coordinates to decrease run time.  Another useful package is nimbleSCR (Bischof et al. 2021) that implements custom sampling distributions to increase sampling speed and efficiency and recent methods of localized activity area approaches (see Milleret et al. 2019) along with other custom functionality. These custom distributions from nimbleSCR can be used within our template models, but the idea with the localSCR package is to provide a more user-friendly means of implementing SCR models in nimble similar to the functionality provided by the jagsUI package (Keller 2019) for JAGS (Plummer 2017). Our intention is to ease the transistion for users that were previously using JAGS or those that have yet to switch over to Bayesian methods because of the complexity of implementing these approaches.

## Installation

You can install the development version of localSCR like so:

```{r }
install_github("sitkensis22/localSCR")
```

## Example

This is a basic example which shows you a workflow for a uniform state-space with a habitat mask:

```{r example}

# load package
library(localSCR)

```

## (1) Simulate a single trap array with random positional noise and create state-space

```{r, fig.align = "center"}

# simulate a single trap array with random positional noise
x <- seq(-800, 800, length.out = 5)
y <- seq(-800, 800, length.out = 5)
traps <- as.matrix(expand.grid(x = x, y = y))
traps <- traps + runif(prod(dim(traps)),-20,20) # add some random noise to locations

sigma = 300 # simulate sigma of 300 m
mycrs = 32608 # EPSG for WGS 84 / UTM zone 8N

# create state-space
Grid = grid_classic(X = traps, crs_ = mycrs, buff = 3*sigma, res = 100)

# make plot of grid and trap locations
par(mfrow=c(1,1))
plot(Grid$grid, pch=19)
points(traps, col="blue",pch=20)


```


## (2) Simulate SCR data and make a plot of it.

```{r, fig.align = "center"}

# simulate SCR data
data3d = sim_classic(X = traps, ext = Grid$ext, crs_ = mycrs, sigma = sigma, prop_sex = 1, N = 200, K = 4, base_encounter = 0.25, enc_dist = "binomial", hab_mask = FALSE, setSeed = 50)

# inspect simulated data
str(data3d)


# make simple plot
par(mfrow=c(1,1))
plot(Grid$grid, pch=20,ylab="Northing",xlab="Easting")
points(traps, col="blue",pch=20)
points(data3d$s,col="red",pch = 20)
points(data3d$s[which(apply(data3d$y,1,sum)!=0),],col="green",pch = 20)


```



# (3) Workflow for simple SCR model with sex-specific sigma, binomial encounter distribution, and habitat mask

```{r, fig.align = "center"}

# simulate a single trap array with random positional noise
x <- seq(-800, 800, length.out = 5)
y <- seq(-800, 800, length.out = 5)
traps <- as.matrix(expand.grid(x = x, y = y))
traps <- traps + runif(prod(dim(traps)),-20,20) # add some random noise to locations

sigma = c(220, 300) # simulate sex-specific
mycrs = 32608 # EPSG for WGS 84 / UTM zone 8N
pixelWidth = 100 # store pixelWidth

# create state-space grid and extent
Grid = grid_classic(X = traps, crs_ = mycrs, buff = 3*max(sigma), res = pixelWidth) 

# create polygon for mask
poly = sf::st_sfc(sf::st_polygon(x=list(matrix(c(-1765,-1765,1730,-1650,1600,1650,0,1350,-800,1700,-1850,1000,-1765,-1765),ncol=2, byrow=TRUE))), crs =  mycrs)

# create habitat mask
hab_mask = mask_polygon(poly = poly, grid = Grid$grid, crs_ = mycrs, prev_mask = NULL)

# simulate data for uniform state-space and habitat mask
data3d = sim_classic(X = traps, ext = Grid$ext, crs_ = mycrs, sigma = sigma, prop_sex = 0.7, N = 200, K = 4, base_encounter = 0.15, enc_dist = "binomial",hab_mask = hab_mask, setSeed = 50)

# get initial activity center starting values
s.st3d = initialize_classic(y=data3d$y, M=400, X=traps, buff = 3*max(sigma), hab_mask = hab_mask)

# make simple plot
par(mfrow=c(1,1))
plot(Grid$grid, pch=20, xlim=c(-2000,2000),ylim=c(-2000,2000))
points(traps, col="blue",pch=20)
plot(poly, add=TRUE)
points(s.st3d,col="red",pch=20)

# rescale inputs
rescale_list = rescale_classic(X = traps, ext = Grid$ext, s.st = s.st3d, hab_mask = hab_mask)

# store rescaled extent
ext = rescale_list$ext

# prepare data
data = list(y=data3d$y)
data$y = data$y[which(apply(data$y, 1, sum)!=0),,] # remove augmented records 
data$y = apply(data$y, c(1,2), sum) # covert to 2d by summing over occasions

# add rescaled traps
data$X = rescale_list$X

# prepare constants (note get density in km2 rather than m2)
constants = list(M = 400,n0 = nrow(data$y),J=dim(data$y)[2], K=dim(data3d$y)[3],x_lower = ext[1], x_upper = ext[2], y_lower = ext[3], y_upper = ext[4],sigma_upper = 1000, A = (sum(hab_mask)*(pixelWidth/1000)^2),pixelWidth=pixelWidth)

# augment sex
data$sex = c(data3d$sex,rep(NA,constants$M-length(data3d$sex)))

# add z and zeros vector data for latent inclusion indicator
data$z = c(rep(1,constants$n0),rep(NA,constants$M - constants$n0))
data$zeros =  c(rep(NA,constants$n0),rep(0,constants$M - constants$n0))

# add hab_mask and OK for habitat check
data$hab_mask = hab_mask
data$OK = rep(1,constants$M)

# get initial activity center starting values
s.st3d = rescale_list$s.st

# define all initial values
inits = list(sigma = runif(2, 250, 350), s = s.st3d,psi=runif(1,0.2,0.3), p0 = runif(1, 0.05, 0.15),pOK = data$OK,z=c(rep(NA,constants$n0),rep(0,constants$M-constants$n0)))

# parameters to monitor
params = c("sigma","psi","p0","N","D","psi_sex")

# get model
scr_model = get_classic(dim_y = 2, enc_dist = "binomial",sex_sigma = TRUE,hab_mask=TRUE)

# run model
tictoc::tic() # track time elapsed
out = run_classic(model = scr_model, data=data, constants=constants, inits=inits, params = params,niter = 8000, nburnin=1000, thin=1, nchains=2, parallel=TRUE, RNGseed = 500)
tictoc::toc()

# summarize output
samples = do.call(rbind, out)
par(mfrow=c(1,1))
hist(samples[,which(dimnames(out[[1]])[[2]]=="N")], xlab = "Abundance", xlim = c(0,500), main="")
abline(v=200, col="red") # add line for simulated abundance

# summarize MCMC samples
nimSummary(out)

```
